RSA Project Report Paper

### 2016-11-29 Paper/Paper.tex

	\indent A Lucas sequence is a recurrence relation that satisfies of $x^2 - Px+Q$ where with P and Q being integers.In the paper there are seven different identities for Lucas sequences described.These identities stem from this first equation (which is noted by the paper to be the formal definition of a Lucas Sequence)  $U_n(P,Q) = (\alpha^{n}-\beta^{n})/(\alpha-\beta), V_n(P,Q) = \alpha^{n}+\beta^{n}$ where $\alpha$ and $\beta$  are the roots of $x^{}2 - Px+Q$. All the identities won't be outlined here, however there are multiple identities used in the algorithm for the P+1 method. 
	\\\\
	\indent The algorithm itself as outlined in the paper is similar to the following. Let $p = (\prod^{k}_{i=1}q^{\alpha_i}_i - 1)$ where p is a prime factor of N. If we define R in the same way as in the P-1 method then $p+1|R$ and $p|U_R(P,Q)$ thus $p|(U_R(P,Q),N)$. The issues with this method is the calculation of $U_R$ which can be a very large number as the size of R may be very large. A method is described to get around this which ends in an assumption that Q may be set to be equal to one in the calculations of all Lucas sequences involved in this algorithm. 
	\\\\
	\indent There is a second bypass involved to decrease the computational load that leads to the primary implementation done for our assignment. The details to this implementation are outlined in the paper itself. This method was implemented on a laptop computer with a I3 Intel processor that clocks at 2.2 Gigahertz with a RAM size of 4.0 gigabytes. Firstly define a Lucas Sequence to be $V_n = V_{n-1}+V{n_2}$ where $V_0 = 2$ and $V_1 = a$ with a being an integer larger than 2. Then create a loop that continuously calculates $V_{n!}$ where $n \geq 2$. Then after every iteration of this loop calculate the GCD of (N,$V_{n!}-2$) this was done using the Euclid's Algorithm. When the GCD is not equal to either 1 or N, then you have a non trivial factor of N. The major difficulty of implementing this algorithm was the computational overload of calculating subsequent factorial Lucas Sequences. There was a small work around to this to relieve some of the burden of this by recognizing that $V_{n!}(a) = V_n({V_{n-1}})$ meaning if we allow the input to the Lucas sequence to be the output to the previous Lucas Sequence, we can get the factorials of the initial input a without having to actually calculate subsequent factorials of that Lucas Sequence. Without this work around even five iterations of this loop quickly became lengthy to calculate, as it would be $V_120$, an extreme amount of stack calls on the limited hardware being used. The numbers still become prohibitively large for the hardware fairly quickly, in about 15 iterations. So a limit on the number of iterations was engraved in the method (about 15 or so). The paper itself mentions in the implementation and results section that the P+1 method was about 9 times slower than the P-1 method possibly likely due to the calculation of a Lucas sequence on top of the calculation of the factorial that the P-1 method already had. When actually run on the original modulus the algorithm once again found 809 to be a factor about 5 seconds. Use of the method on the modulus with 809 divided out produced no further results, even when run for over 20+ hours. At this time an estimate for the time the program would actually factor the second modulus has not been produced. The only new information that could be extrapolated from this would be that the probability of the p+1 of the larger number being comprised of mostly small factors would be low provided that the premise of the algorithm itself is to be believed. This ends the discussion of William's P+1 method and our attempt to implement it for our assignment.

@@

		Again, to decide whether $\legendre{N}{p} = 1$, we implemented both Euler's Criterion and Legendre Symbol and again probably because our factor base was not that large, we did not notice that much of a difference in the Law of Quadratic Reciprocity over Euler's Criterion. Since Professor Boklan, suggested us to use Law of Quadratic Reciprocity, we ended up using this to test $\legendre{N}{p}$.

		\item \textit{Pre-Sieving:} \itemgap
		\\ In this stage, we did a lot optimizations. First, we use an array as a dictionary, which means we use prime valuse as the indices to the array, meaning since our B is one million, our array length is also one million. The reason why we do this is because we to be extremely fast in our sieving process. Since the array get operation is $\mathcal{O}(1)$ this is very fast. The downside, however, is obviously wasting memory. We then use this approach to save 3 informations. First one, is the roots of quadratic residue. Since every odd prime has two solutions, we use a two dimensional array to save both. Secondly, we also cache the value of $\log p$, since we do not want any unnecessary calculations in our sieving process. Finally, we also store the offset of x from the start position $\lfloor \sqrt{N} \rfloor - M$. We did not see any resource mention this, but we think it is useful, since we can very easily obtain the smooth numbers by just adding the offset to the starting position.

@@

		First, we will explain why finding the null space is equal to finding the sum of rows equal to zero, because initially we did not realize this ourselves from all the resources we read. Recall the how we perform the product of the matrices: for each row multiply a coefficient add rows together. We are working over $\mathbb{F}_2$, the multiplier 0 means not using this row, 1 means using this row, which is why we are checking each element of the vector in the null space; if we found a 1, we find the corresponding row in matrix and get related $x$ and all the prime factors of $Q(x)$.

@@

>>>>

	\indent A Lucas sequence is a recurrence relation that satisfies of $x^2 - Px+Q$ where with P and Q being integers.In the paper there are seven different identities for Lucas sequences described.These identities stem from this first equation (which is noted by the paper to be the formal definition of a Lucas Sequence)  $U_n(P,Q) = (\alpha^{n}-\beta^{n})/(\alpha-\beta), V_n(P,Q) = \alpha^{n}+\beta^{n}$ where $\alpha$ and $\beta$  are the roots of $x^{}2 - Px+Q$. All the identities won't be outlined here, however there are multiple identities used in the algorithm for the P+1 method. 
	\\\\
	\indent The algorithm itself as outlined in the paper is similar to the following. Let $p = (\prod^{k}_{i=1}q^{\alpha_i}_i - 1)$ where p is a prime factor of N. If we define R in the same way as in the P-1 method then $p+1|R$ and $p|U_R(P,Q)$ thus $p|(U_R(P,Q),N)$. The issues with this method is the calculation of $U_R$ which can be a very large number as the size of R may be very large. A method is described to get around this which ends in an assumption that Q may be set to be equal to one in the calculations of all Lucas sequences involved in this algorithm. 
	\\\\
	\indent There is a second bypass involved to decrease the computational load that leads to the primary implementation done for our assignment. The details to this implementation are outlined in the paper itself. This method was implemented on a laptop computer with a I3 Intel processor that clocks at 2.2 Gigahertz with a RAM size of 4.0 gigabytes. Firstly define a Lucas Sequence to be $V_n = V_{n-1}+V{n_2}$ where $V_0 = 2$ and $V_1 = a$ with a being an integer larger than 2. Then create a loop that continuously calculates $V_{n!}$ where $n \geq 2$. Then after every iteration of this loop calculate the GCD of (N,$V_{n!}-2$) this was done using the Euclid's Algorithm. When the GCD is not equal to either 1 or N, then you have a non trivial factor of N. The major difficulty of implementing this algorithm was the computational overload of calculating subsequent factorial Lucas Sequences. There was a small work around to this to relieve some of the burden of this by recognizing that $V_{n!}(a) = V_n({V_{n-1}})$ meaning if we allow the input to the Lucas sequence to be the output to the previous Lucas Sequence, we can get the factorials of the initial input a without having to actually calculate subsequent factorials of that Lucas Sequence. Without this work around even five iterations of this loop quickly became lengthy to calculate, as it would be $V_120$, an extreme amount of stack calls on the limited hardware being used. The numbers still become prohibitively large for the hardware fairly quickly, in about 15 iterations. So a limit on the number of iterations was engraved in the method (about 15 or so). The paper itself mentions in the implementation and results section that the P+1 method was about 9 times slower than the P-1 method possibly likely due to the calculation of a Lucas sequence on top of the calculation of the factorial that the P-1 method already had. When actually run onfthe original modulus the algorithm once again found 809 to be a factor about 5 seconds. Use of the method on the modulus with 809 divided out produced no further results, even when run for over 20+ hours. At this time an estimate for the time the program would actually factor the second modulus has not been produced. The only new information that could be extrapolated from this would be that the probability of the p+1 of the larger number being comprised of mostly small factors would be low provided that the premise of the algorithm itself is to be believed. This ends the discussion of William's P+1 method and our attempt to implement it for our assignment.

@@

		Again, to decide whether $\legendre{N}{p} = 1$, we implemented both Euler's Criterion and Legendre Symbol and again probably because our factor base was not that large, we did not notice that much of a difference in the Law of Quadratic Reciprocity over Euler's Criterion. Since Professor Boklan, suggested us to use the Law of Quadratic Reciprocity, we ended up using this to test $\legendre{N}{p}$.

		\item \textit{Pre-Sieving:} \itemgap
		\\ In this stage, we did a lot optimizations. First, we use an array as a dictionary, which means we use prime valuse as the indices to the array, meaning since our B is one million, our array length is also one million. The reason why we do this is because we to be extremely fast in our sieving process. Since the array get operation is $\mathcal{O}(1)$ this is very fast. The downside, however, is obviously wasting memory. We then use this approach to save three pieces ofinformation.First one, is the roots of quadratic residue. Since every odd prime has two solutions, we use a two dimensional array to save both. Secondly, we also cache the value of $\log p$, since we do not want any unnecessary calculations in our sieving process. Finally, we also store the offset of x from the start position $\lfloor \sqrt{N} \rfloor - M$. We did not see any resource mention this, but we think it is useful, since we can very easily obtain the smooth numbers by just adding the offset to the starting position.

@@

		First, we will explain why finding the null space is equal to finding the sum of rows equal to zero, because initially we did not realize this ourselves from all the resources we read. After finding the null space, each vector in the null space will contain coefficients in which each coefficient corresponds to a row in the original matrix.Recall we perform the product of the matrices: for each row multiply by a coefficient, and add the addows together. We are working over $\mathbb{F}_2$, the multiplier 0 means not using this row, 1 means using this row, which is why we are checking each element of the vector in the null space; if we found a 1, we find the corresponding row in matrix and get related $x$ and all the prime factors of $Q(x)$.

@@

		\begin{figure}[h]

			\includegraphics[width=0.6\textwidth]{tree.png}
		\end{figure}
